The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2023-05-14 MS_ToSubmit/MS_Burdwood_foodweb.tex
Co-authors order
Lovrich & Delpiani suggs

    \author[]{Fabiana L. Capitanio\textsuperscript{f}%
  %
  }

    \author[]{Sergio M. Delpiani\textsuperscript{d}%
  %
  }

    \author[]{Juan Martín Díaz de Astarloa\textsuperscript{d}%
  %
  }

@@

    \author[]{Andrea Raya Rey\textsuperscript{a,c,e}%
  %
  }

    \author[]{Mariela L. Spinelli\textsuperscript{f}%
  %
  }

    \author[]{Diego Urteaga\textsuperscript{g}%
  %
  }

    \author[]{Virginia A. García Alonso\textsuperscript{f}%
  %
  }

@@

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \affiliation[3]{Instituto de Ciencias Polares, Ambiente y Recursos
Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Argentina}
    \affiliation[4]{Instituto de Investigaciones Marinas y Costeras
(IIMYC), Universidad Nacional de Mar del Plata-CONICET, Argentina}
    \affiliation[5]{Wildlife Conservation Society, Argentina}
    \affiliation[6]{Instituto de Biodiversidad y Biologia Experimental y
Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, Argentina}
    \affiliation[7]{Museo Argentino de Ciencias Naturales ``Bernardino
Rivadavia'', Argentina}
    \affiliation[8]{Instituto de Diversidad y Ecologia Animal
(IDEA-CONICET)}
    \affiliation[9]{Facultad de Ciencias Exactas, Fisicas y Naturales,
Universidad Nacional de Cordoba (UNC), Argentina}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá - Banco
  Burdwood I and II (MPA N-BB, 53º--55ºS and 56º--62ºW, respectively),
  which are associated to a submarine plateau and its southern adjacent
  deep slope region, respectively. Here, we present the first detailed
  analysis of the network of predator-prey interactions (food web) for
  the MPA N-BB ecosystem. We applied a network approach to characterise
  the food web in terms of complexity and structure, and identifying the
  species' role in such a framework. In terms of complexity, the MPA
  N-BB food web consisted of 1780 predator-prey interactions and 375
  species, with a link density of 4.75 and a connectance of 0.01. In
  terms of structure, almost half of the consumers were omnivores
  (0.48), and the network displayed a small-world pattern. These
  findings suggest that the ecosystem might be vulnerable to external
  perturbations targeting highly connected species, although structural
  properties might provide resilience and resistance, resulting in a
  rearranged structure that preserves its original functions.
  Furthermore, we identified several species as important in terms of
  different aspects of trophic structure and functioning, and response
  to perturbations. We suggest that generalist species, mainly fishes,
  play a crucial role in the ecosystem's bentho-pelagic coupling process
  and that other species, besides the longtail southern cod
  \emph{Patagonotothen ramsayi} and the Fueguian sprat \emph{Sprattus
  fuegensis}, should be considered as relevant energy transfers for the
  ecosystem. Finally, we argue that the diversity of species, including
  both the benthic and pelagic habitats, is responsible for seccuring
  the connectivity within the food web against perturbations, therefore
  contributing to the structure and stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Marine Protected
Area \sep 
    Southwest Atlantic
  \end{keyword}

@@

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB, 53º--55ºS 56º--62ºW), which is the focus of
this work, is unique since these MPAs are associated to a submarine
plateau and its southern adjacent deep slope region, respectively
(Falabella, 2017). In addition, such MPAs are part of a network of
protected areas in the sub-Antarctic area (jointly with MPA Yaganes)
that aims to protect and strengthen the functioning of this southern
region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; Schejter \& Bremec, 2019; Sirito, 2019), and has an important role
in the life history of fishes as a food source, refuge and nursery area
(Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al., 2022;
Florencia et al., 2023; García Alonso et al., 2018; Troccoli et al.,
2020). The maintenance of this singular community is related to local
and regional oceanographic processes, including the circulation of the
rich Malvinas (Falkland) current in the area (Guerrero et al., 1999; A.
R. Piola \& Gordon, 1989) and the upwelling and mixing phenomena (Matano
et al., 2019). The input of nutrients from the Malvinas (Falkland)
current also supports a diverse plankton community (Guinder et al.,
2020). Overall, 811 benthic and plankton species have been identified
for the MPA N-BB ecosystem, where 349 were reported for the first time
in the area in recent years (APN, 2022). However, identifying the main
species involved in the maintenance of ecosystem services and health as
well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure, meaning a shorter food-chain length and a greater trophic
overlap and redundancy, than other sub-Antarctic areas, such as the
Atlantic coast of Tierra del Fuego. The Fuegian sprat \emph{Sprattus
fuegensis} and longtail southern cod \emph{Patagonotothen ramsayi} are
considered the most plausible `wasp-waist' species (Riccialdelli et al.,
2020).

High-latitude marine ecosystems, such as the MPA N-BB, are considered
complex systems in terms of biodiversity and ecological interactions
(Cordone et al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan
et al., 2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020; Schejter et al.,
2016, 2020), a better understanding of species interactions' complexity
and structure is needed. This aspect can be tackled by analysing one of
the most frequent relationships between species: the predator-prey
interaction (Bascompte, 2009). The sum of predator-prey or trophic
interactions of a particular region is referred to as a food web,
representing the roadmap for matter and energy flow in an ecosystem. In
recent years, network approaches have been successfully applied to study
complex high-latitude marine ecosystems, improving our knowledge on
structure, functioning, and response to environmental/anthropogenic
changes (Cordone et al., 2018; Funes et al., 2022; Kortsch et al., 2015;
Marina et al., 2023). Among anthropogenic threats, it is worth
mentioning that contaminants like mercury and microplastics have been
recently reported as important threats to the MPA N-BB region (Cossi et
al., 2021; Di Mauro et al., 2022; Fioramonti et al., 2022).

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 34,000 km\textsuperscript{2} circumscribed
by the 200 m isobath, between 54º--55ºS and 56º--62ºW, with a slight
slope extended nearly 370 km east--west. Physical features in the BB are
fairly stable, with salinity averaging 34 all year round and temperature
ranging between 4 and 8ºC (Acha et al., 2004; Guerrero et al., 1999; A.
Piola \& Falabella, 2009). The BB is surrounded by steep flanks of up to
4000 m depth through which strong currents circulate (Matano et al.,
2019; A. R. Piola \& Gordon, 1989; Reta, 2014). Such deep slope forms
the basis of N-BB II, protecting about 32,000 km\textsuperscript{2}.
Intense upwelling and mixing occur over it, entraining deep
nutrient-rich waters into the photic layer (Matano et al., 2019; A.
Piola \& Falabella, 2009) and resulting in a fairly homogeneous water
column both spatially and temporally (Glorioso \& Flather, 1995;
Guerrero et al., 1999; Matano et al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in APN, 2022), a joint management plan was recently proposed
(APN, 2022). This is why, the study area of the present work includes
both MPAs.

\begin{figure}
\includegraphics[width=1\linewidth]{MPABurdwood_map} \caption{Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve.}\label{fig:figure1}
\end{figure}

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to 16 research cruises conducted
in the MPAs N-BB I and II during 2014-2019. Here it is important to note
that the sampling effort was greater in the MPA N-BB I. Furthermore, we
took into account personal communications from experts belonging to the
working group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina, Saravia, et al. (2018).

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web
(N-BB I and II) consisted of 1780 predator-prey interactions and 375
species, where 93\% of them were defined at the species taxonomical
level (Figure 2, Table S2). The food web presented a link density (e.g.,
the average number of interactions per species) of 4.75, and a
connectance of 0.01. Almost half of the consumers were omnivores (0.48),
feeding on sources at different trophic levels. The food web displayed a
small-world pattern, meaning that the path length was lower and the
clustering coefficient higher than the random networks (Table 2).

@@

\caption{Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level.}\label{fig:figure2}
\end{figure}

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
375 & 1780 & 4.75 & 0.01 & 0.48 & 2.96 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

@@

\hypertarget{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}{%
\subsection{The food web of the MPA Namuncurá - Banco Burdwood
ecosystem}\label{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}}

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina, Saravia, et al., 2018). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be five times lower than that of Beagle Channel, an
adjacent coastal area (Rodriguez et al., 2022).

@@

In summary, the food web of the MPA N-BB presents a combination of
network properties that makes it unique in terms of network resolution,
complexity, and structural pattern. All this suggests that the ecosystem
might be fragile to external perturbations targeting highly connected
species. However, structural properties might provide resilience and
resistance with the final outcome of a rearranged structure maintaining
its functions.

@@

We have identified that most of the consumers in the food web of the MPA
N-BB either have a narrow diet or are specialists, while few present a
broad diet. The most evident generalist species are \emph{Patagonotothen
guntheri} (Covatti Ale et al., 2022), \emph{P. ramsayi} (Fischer et al.,
2022), \emph{Dissostichus eleginoides} (Troccoli et al., 2020),
\emph{Bathyraja brachyurops} (Belleggia et al., 2008), and \emph{B.
griseocauda} (Bellegia et al., 2014), with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4),
acting as predator and prey, they might be important links between lower
and higher trophic levels. This result is in agreement with the sole
analysis, using stable isotopes, that exists so far for the trophic
structure of the MPA N-BB (Riccialdelli et al., 2020), and resembles
other high-latitude marine systems of the Southwest Atlantic and
Antarctic regions (Arkhipkin \& Laptikhovsky, 2013; Marina, Salinas, et
al., 2018). The importance of these particular generalist species also
arises since they feed on the benthic and pelagic habitats (Covatti Ale
et al., 2022; Fischer et al., 2022; Troccoli et al., 2020), linking
these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators
in the food web of the MPA N-BB. This suggests that there are dominant
food sources from which most consumers depend on and from where the
ecosystem energy is being transferred to the upper trophic levels. The
most demanded source we identified in this study (i.e.~detritus)
supports the abundant benthic community of filter-feeders (Schejter et
al., 2016). Furthermore, we found that the second and third-most
consumed prey were diatoms and species of \emph{Euphausia},
respectively, which are essential sources for the diverse zooplankton
community (Spinelli et al., 2020), mid-TL consumers like the Fuegian
sprat \emph{Sprattus fuegensis} (Padovani et al., 2021) and
\emph{Patagonotothen ramsayi} (Fischer et al., 2022), and top predators
such as the black-browed and grey-headed albatrosses (\emph{Thalassarche
melanophris} and \emph{Thalassarche chrysostoma}, respectively) (Catry
et al., 2004), and baleen whales (species of the genera Balaenoptera and
Eubalaena) (Valenzuela et al., 2018).

@@

The food web studied in the present work might be more representative of
the MPA Namuncurá - Banco Burdwood I ecosystem, which comprises the
shallow submarine plateau called Burdwood Bank. This is related to the
sampling effort that was conducted during the research cruises in the
former MPA compared to the MPA N-BB II (i.e.~deep flanks to the south).
As a consequence, most of the data we used to build the network come
from studies performed in the MPA N-BB I. Despite this fact, we decided
to build the food web considering both MPAs due to the tight
oceanographic and ecological connection that exists among them (APN,
2022 and references therein).

>>>>

@@

@@

    \author[]{Fabiana L. Capitanio\textsuperscript{d,e}%
  %
  }

    \author[]{Sergio M. Delpiani\textsuperscript{f}%
  %
  }

    \author[]{Juan Martín Díaz de Astarloa\textsuperscript{e}%
  %
  }

    \author[]{Virginia A. García Alonso\textsuperscript{d,e}%
  %
  }

    \author[]{Andrea Raya Rey\textsuperscript{a,c,g}%
  %
  }

    \author[]{Mariela L. Spinelli\textsuperscript{d,e}%
  %
  }

    \author[]{Diego Urteaga\textsuperscript{j}%
  %
  }

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \affiliation[3]{Instituto de Ciencias Polares, Ambiente y Recursos
Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Argentina}
    \affiliation[4]{Facultad de Ciencias Exactas y Naturales,
Universidad de Buenos Aires (UBA), Argentina}
    \affiliation[5]{Instituto de Biodiversidad y Biologia Experimental y
Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, Argentina}
    \affiliation[6]{Instituto de Investigaciones Marinas y Costeras
(IIMYC), Universidad Nacional de Mar del Plata-CONICET, Argentina}
    \affiliation[7]{Wildlife Conservation Society, Argentina}
    \affiliation[8]{Facultad de Ciencias Exactas, Fisicas y Naturales,
Universidad Nacional de Cordoba (UNC), Argentina}
    \affiliation[9]{Instituto de Diversidad y Ecologia Animal
(IDEA-CONICET)}
    \affiliation[10]{Museo Argentino de Ciencias Naturales ``Bernardino
Rivadavia'', Argentina}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá -
  Burdwood Bank I and II (MPA N-BB, 53º--55ºS and 56º--62ºW,
  respectively), which are associated to a submarine plateau and its
  southern adjacent deep slope region, respectively. Here, we present
  the first detailed analysis of the network of predator-prey
  interactions (food web) for the MPA N-BB ecosystem. We applied a
  network approach to characterise the food web in terms of complexity
  and structure, and identifying the species' role in such a framework.
  In terms of complexity, the MPA N-BB food web consisted of 1780
  predator-prey interactions and 375 species, with a link density of
  4.75 and a connectance of 0.01. In terms of structure, almost half of
  the consumers were omnivores (0.48), and the network displayed a
  small-world pattern. These findings suggest that the ecosystem might
  be vulnerable to external perturbations targeting highly connected
  species, although structural properties might provide resilience and
  resistance, resulting in a rearranged structure that preserves its
  original functions. Furthermore, we identified several species as
  important in terms of different aspects of trophic structure and
  functioning, and response to perturbations. We suggest that generalist
  species, mainly fishes, play a crucial role in the ecosystem's
  bentho-pelagic coupling process and that other species, besides the
  longtail southern cod \emph{Patagonotothen ramsayi} and the Fueguian
  sprat \emph{Sprattus fuegensis}, should be considered as relevant
  energy transfers for the ecosystem. Finally, we argue that the
  diversity of species, including both the benthic and pelagic habitats,
  is responsible for seccuring the connectivity within the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Burdwood Bank \sep 
    Southwest Atlantic
  \end{keyword}

@@

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá -
Burdwood Bank I and II (MPA N-BB, 53º--55ºS 56º--62ºW), which is the
focus of this work, is unique since these MPAs are associated to a
submarine plateau and its southern adjacent deep slope region,
respectively (Falabella, 2017). In addition, such MPAs are part of a
network of protected areas in the sub-Antarctic area (jointly with MPA
Yaganes) that aims to protect and strengthen the functioning of this
southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; Schejter \& Bremec, 2019; Sirito, 2019), and has an important role
in the life history of fishes as a food source, refuge and nursery area
(Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al., 2022;
Florencia et al., 2023; García Alonso et al., 2018; Troccoli et al.,
2020; Vazquez et al., 2018). The maintenance of this singular community
is related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; Piola \& Gordon, 1989) and the upwelling and
mixing phenomena (Matano et al., 2019). The input of nutrients from the
Malvinas (Falkland) current also supports a diverse plankton community
(Guinder et al., 2020). Overall, 811 benthic and plankton species have
been identified for the MPA N-BB ecosystem, where 349 were reported for
the first time in the area in recent years (Administracion de Parques
Nacionales, 2022). However, identifying the main species involved in the
maintenance of ecosystem services and health as well as for management
and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure, meaning a shorter food-chain length and a greater trophic
overlap and redundancy, than other sub-Antarctic areas, such as the
continental shelf off Tierra del Fuego. The Fuegian sprat \emph{Sprattus
fuegensis} and longtail southern cod \emph{Patagonotothen ramsayi} are
considered the most plausible `wasp-waist' species (Riccialdelli et al.,
2020).

High-latitude marine ecosystems, such as the MPA N-BB, are considered
complex systems in terms of biodiversity and ecological interactions
(Cordone et al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan
et al., 2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (Administracion de Parques Nacionales, 2022; Guinder
et al., 2020; Schejter et al., 2016, 2020), a better understanding of
species interactions' complexity and structure is needed. This aspect
can be tackled by analysing one of the most frequent relationships
between species: the predator-prey interaction (Bascompte, 2009). The
sum of predator-prey or trophic interactions of a particular region is
referred to as a food web, representing the roadmap for matter and
energy flow in an ecosystem. In recent years, network approaches have
been successfully applied to study complex high-latitude marine
ecosystems, improving our knowledge on structure, functioning, and
response to environmental/anthropogenic changes (Cordone et al., 2018;
Funes et al., 2022; Kortsch et al., 2015; Marina et al., 2023). Among
anthropogenic threats, it is worth mentioning that contaminants like
mercury and microplastics have been recently reported as important
threats to the MPA N-BB region (Cossi et al., 2021; Di Mauro et al.,
2022; Fioramonti et al., 2022); also fishing vessels are allowed to
operate in the western section of the MPA N-BB (i.e.~Marine National
Reserve category), altering the stocks of commercially important fish
species (Administracion de Parques Nacionales, 2022; Martínez et al.,
2021). Moreover, there is a potential hazard related to the effects of
offshore activities (exploration and explotation) to the west of the MPA
N-BB (Administracion de Parques Nacionales, 2022).

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws
26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau
called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth,
N-BB I and N-BB II, respectively (Administracion de Parques Nacionales,
2023; Tombesi et al., 2020) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 34,000 km\textsuperscript{2} circumscribed
by the 200 m isobath, between 54º--55ºS and 56º--62ºW, with a slight
slope extended nearly 370 km east--west. Physical features in the BB are
fairly stable, with salinity averaging 34 all year round and temperature
ranging between 4 and 8ºC (Acha et al., 2004; Guerrero et al., 1999;
Piola \& Falabella, 2009). The BB is surrounded by steep flanks of up to
4000 m depth through which strong currents circulate (Matano et al.,
2019; Piola \& Gordon, 1989; Reta, 2014). The N-BB II includes such a
deep slope, protecting about 32,000 km\textsuperscript{2}. Intense
upwelling and mixing occur over it, entraining deep nutrient-rich waters
into the photic layer (Matano et al., 2019; Piola \& Falabella, 2009)
and resulting in a fairly homogeneous water column both spatially and
temporally (Glorioso \& Flather, 1995; Guerrero et al., 1999; Matano et
al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in Administracion de Parques Nacionales, 2022), a joint
management plan was recently proposed (Administracion de Parques
Nacionales, 2022). This is why, the study area of the present work
includes both MPAs.

\begin{figure}
\includegraphics[width=1\linewidth]{MPABurdwood_map} \caption{Marine Protected Areas Namuncurá - Burdwood Bank I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve.}\label{fig:figure1}
\end{figure}

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to 16 research cruises conducted
in the MPAs N-BB I and II during 2014-2019. It is noteworthy that the
sampling effort was greater in the MPA N-BB I. Furthermore, we took into
account personal communications from experts belonging to the working
group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbour nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina, Saravia, et al. (2018).

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web (N-BB
I and II) consisted of 1780 predator-prey interactions and 375 species,
where 93\% of them were defined at the species taxonomical level (Figure
2, Table S2). The food web presented a link density (e.g., the average
number of interactions per species) of 4.75, and a connectance of 0.01.
Almost half of the consumers were omnivores (0.48), feeding on sources
at different trophic levels. The food web displayed a small-world
pattern, meaning that the path length was lower and the clustering
coefficient higher than the random networks (Table 2).

@@

\caption{Graph of the food web for the MPA Namuncurá - Burdwood Bank. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level.}\label{fig:figure2}
\end{figure}

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Burdwood Bank food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
375 & 1780 & 4.75 & 0.01 & 0.48 & 2.96 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

@@

\hypertarget{the-food-web-of-the-mpa-namuncuruxe1---burdwood-bank-ecosystem}{%
\subsection{The food web of the MPA Namuncurá - Burdwood Bank
ecosystem}\label{the-food-web-of-the-mpa-namuncuruxe1---burdwood-bank-ecosystem}}

@@

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina, Saravia, et al., 2018). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Burdwood Bank (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to much lower than that of Beagle Channel (0.05), an adjacent
coastal area (Rodriguez et al., 2022).

@@

In summary, the food web of the MPA N-BB presents a combination of
network properties that makes it unique in terms of network resolution,
complexity, and structural pattern. All this suggests that the ecosystem
might be fragile to external perturbations targeting highly connected
species, which in turn coincides to be commercial exploited species as
fishes (Martínez et al., 2015). However, structural properties might
provide resilience and resistance with the final outcome of a rearranged
structure maintaining its functions.

@@

We have identified that most of the consumers in the food web of the MPA
N-BB either have a narrow diet or are specialists, while few present a
broad diet. The most evident generalist species are \emph{Patagonotothen
guntheri} (Covatti Ale et al., 2022), \emph{P. ramsayi} (Fischer et al.,
2022), \emph{Dissostichus eleginoides} (Troccoli et al., 2020),
\emph{Bathyraja brachyurops} (Belleggia et al., 2008), and \emph{B.
griseocauda} (Bellegia et al., 2014), with more than 25 potential prey.
Since these species present mid-trophic positions in the food web (3-4),
acting as predator and prey, they might be important links between lower
and higher trophic levels. This result is in agreement with the sole
analysis, using stable isotopes, that exists so far for the trophic
structure of the MPA N-BB (Riccialdelli et al., 2020), and resembles
other high-latitude marine systems of the Southwest Atlantic and
Antarctic regions (Arkhipkin \& Laptikhovsky, 2013; Marina, Salinas, et
al., 2018). The importance of these particular generalist species also
arises since they feed in the benthic and pelagic habitats (Covatti Ale
et al., 2022; Fischer et al., 2022; Troccoli et al., 2020), linking
these realms and contributing to the vertical carbon flow.

On the other hand, a low number of prey are consumed by many predators
in the food web of the MPA N-BB. This suggests that there are dominant
food sources from which most consumers depend on and from where the
ecosystem energy is being transferred to the upper trophic levels. The
most demanded source we identified in this study (i.e.~detritus)
supports the abundant benthic community of filter-feeders (Schejter et
al., 2016), likely feeding on detritus that is constantly resuspended
from the bottom (Martin \& Flores Melo, 2021). Furthermore, we found
that the second and third-most consumed prey were diatoms and species of
\emph{Euphausia}, respectively, which are essential sources for the
diverse zooplankton community (Spinelli et al., 2020), mid-TL consumers
like the Fuegian sprat \emph{Sprattus fuegensis} (Padovani et al., 2021)
and \emph{Patagonotothen ramsayi} (Fischer et al., 2022), and top
predators such as the black-browed and grey-headed albatrosses
(\emph{Thalassarche melanophris} and \emph{Thalassarche chrysostoma},
respectively) (Catry et al., 2004), and baleen whales (species of the
genera Balaenoptera and Eubalaena) (Valenzuela et al., 2018).

@@

The food web studied in the present work might be more representative of
the shallow ecosystem of the submarine plateau called Burdwood Bank, on
which most of the research was focused as the MPA N-BB I was first
created. This is related to the sampling effort that was conducted
during the research cruises in the former MPA compared to the MPA N-BB
II (i.e.~deep flanks to the south). As a consequence, most of the data
we used to build the network come from studies performed in the MPA N-BB
I. Despite this fact, we decided to build the food web considering both
MPAs due to the tight oceanographic and ecological connection that
exists among them (Administracion de Parques Nacionales, 2022 and
references therein).
